---
title: "Correct R package structure"
author: "Kamil Foltyński"
date: "03/04/2019"
output:
  html_document:
    fig_caption: yes
    keep_md: no
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
runtime: shiny
---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = ">")
```

# `testthat` - R package for unit testing

Package `testthat` provides framework and multiple functions for unit testing. To start using it 
we need following R pacakge structure:


```bash
spendworx
├── DESCRIPTION
├── NAMESPACE
├── R
├── inst
├── man
├── spendworx.Rproj
└── tests
    ├── testthat # <-- tests goes there
    │   └── test_sankey_plot.R
    └── testthat.R
```

where `tests` directory and its content can be created using `devtools::use_testthat()` if not exists. 
Test script names must start with **test**.

Test files should contain `testthat::context` which is short decription of its content. 
To create test we use `testthat::test_that` function.

Example test:

```{r}
testthat::context("String length")

testthat::test_that("str_length is number of characters", {
  testthat::expect_equal(stringr::str_length("a"), 1)
  testthat::expect_equal(stringr::str_length("ab"), 2)
  testthat::expect_equal(stringr::str_length("abc"), 3)
})
```

## How to test R package

* All tests in package: `devtools::test("unitTestsWorkshopPackage/")`
* Single test file: `testthat::test_file("unitTestsWorkshopPackage/tests/testthat/test_unify_case_type.R")`
* 

# Make life easier in RStudio

Shortcuts:

Run tests: `Ctrl+Shift+T` (Windows/Linux) or `Cmd+Shift+T` (Mac)

Typical RStudio workflow:

* Write function
* If reference data from file is used then rebuild package (`Ctrl+Shift+B`), 
see more in [reference data part](04-reference-data.Rmd)
* Run tests: `Ctrl+Shift+T`
  * if tests passed then continue
  * else make fixes in function and run tests `Ctrl+Shift+T`

...and repeat until all unit tests work.

# `testthat` functions

* `testthat::expect_true(object/contition)`
* `testthat::expect_equal(object, expected)`
* `testthat::expect_indentical(object, expected)`

```{r, error=TRUE}
testthat::test_that("iris as data.table is equal", {

  dt_iris <- data.table::data.table(iris)
  
  testthat::expect_identical(dt_iris, data.table::data.table(iris))
  testthat::expect_equal(dt_iris, data.table::data.table(iris))
  
})
```
  
* `testthat::skip_if`, `testthat::skip_if_not`
  * tests can be skipped, for example some tests may not be valid on specific envornment:
    * **example #1**: if we running code locally we may not have access to production DB and we use local one which has different type, e.g. is containerized
    * **example #2**: locally LDAP authetntication doesn't work, so we can't test reading login anywhere else than on server
* `testthat::capture_messages` - log messages or meassages printed to terminal can also be tested

# Practice

## Task #1

* clone example repo `https://github.com/Tazovsky/unit-tests-workshop`
* create `testthat` structure

## Task #2

* Test log output of function `unitTestsWorkshopPackage::unify_case_type`

## Task #3

Use `testthat::expect_equal` and `testthat::expect_identical` to test 
function `unitTestsWorkshopPackage::coerce2data.table`, **but** save expected data to `rds` file
and compare it only after reading `readRDS`.

# Sources

* http://r-pkgs.had.co.nz/tests.html
