---
title: "Reference data"
author: "Kamil Foltyński"
date: "03/04/2019"
output:
  html_document:
    fig_caption: yes
    keep_md: no
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
runtime: shiny
---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = ">")
```



# What is reference data? And why should we care?

Along with development process, code changes, but its behaviour and results should not - 
and if they must change, let them change into a way we control. 

Reference data is used as input for unit tests. It's frozen, so it ensures **reproducibility** of results.
During development process **code changes**, but thanks to **reproducibility** it is possible to run 
code (with reference data as input) and **always get same, unchanged results**. However if results vary, then
it's a sign something went wrong and code is broken.

# How to prepare reference data?

* `dput` - the easiest way to prepare reference data is to use `dput` function. It prints object's structure to terminal, so
it is just about copy-pasting from ther to our test script, see example:

```{r}
library(dplyr)
result <- band_members %>% inner_join(band_instruments)
dput(result)
```

so unit test can be witten as:

```{r}
library(dplyr)
ref_data <- structure(
  list(
    name = c("John", "Paul"),
    band = c("Beatles",
             "Beatles"),
    plays = c("guitar", "bass")
  ),
  row.names = c(NA, -2L),
  class = c("tbl_df", "tbl", "data.frame")
)

testthat::expect_equal(band_members %>% inner_join(band_instruments), ref_data)
testthat::expect_identical(band_members %>% inner_join(band_instruments), ref_data)
```

* `.rds`/`.RData` - `dput` is useful form small _handy_ data, but imagine reference `data.frame` counting thousands of rows
and pasted into test script...


  <br>
  <details>
  <summary>Click to expand</summary>
  ```{r}
  dput(iris)
  ```
  </details>
  <br>

  In such cases it's worth keeping reference data in R formats such as `.rds`/`.RData`.


* Why should we care?
* How to prepare reference data
  * `dput`, `.Rdata`/`.RDS`, `.png` (because it's lossless)
  * remember about `set.seed` when sampling some data (--> plotly task)
  * `system.file`



